---
title: "LakeN_for_Tami"
author: "Yoav Ben Dor"
date: "6.12.2016"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE ,include=FALSE , root.dir="F:/MyDocuments/GIT/LakeN", out.width = '600px', out.height = '300px'}
knitr::opts_chunk$set(echo = FALSE , warning=FALSE, message=FALSE)

# at home
#setwd("F:/MyDocuments/GIT/LakeN/R")

# at HUJI
setwd("C:/Users/user/Documents/LakeN/R")

source('depthinterp.R')
source('realdepth.R')
source('conctoquant.R')
source('daytoseason.R')
source('datesegmenter.R')
source('datedifference.R')
source('datedifferencedf.R')
source('yearsegmenter.R')
source('yearassigner.R')
source('fitequation.R')
source('datesplitter.R')
source('yearbyelement.R')
source('findoxycline.R')
source('lakemixer.R')
source('NSE.R')
source('lakeNH4.R')


# at home
#setwd("F:/MyDocuments/GIT/LakeN")
# at HUJI
setwd("C:/Users/user/Documents/LakeN")

```


```{r data read, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}
# load relevant libraries
library(plyr)
library(zoo)
library(dplyr)
library(LakeN)
library(plot3D)
library(ggplot2)
library(png)
library(reshape2)

# read the data for the model
# use the read.table function in the cas of a tab=delimited text file
# use the read.csv function for a csv

# data1 - measured gecochmical profiles
data1=read.table("LakeNData1.txt", header=TRUE)

# data2 - lake hypsometric curve including depth to area transformation
data2=read.table("LakeNData2.txt", header=TRUE)

# data3 - lake level measurments for depth corrections
data3=read.table("LakeNData3.txt", header=TRUE)

# data1 - temperature profile measurments
data4=read.table("LakeNData4.txt", header=TRUE)

# split the data set by measurment date
data1.split.orig=split(data1, f=data1$Date)

# fill NA cells with linear interpolation in original data
# initilize list to be filled
data1.split=vector("list", length(data1.split.orig))

first.data.col=3
dates=as.Date(unique(data1$Date, incomparables = FALSE))

for (ii in 1:length(data1.split.orig)){
  temp.df=data1.split.orig[[ii]]
  
  #find columns that are all NA and zero fill them prior to interpolation
  zero.index=NA
  zero.index=which(sapply(temp.df, function(x)all(is.na(x))))
  if (is.integer(zero.index)){
    temp.df[,zero.index]=c(rep.int(0, times=nrow(temp.df)))}
  
  # fill NA values confined between numbers using linear interoplation
  for (jj in first.data.col:ncol(temp.df)){
    app.var=as.double(na.approx(x=temp.df$Depth,as.numeric(unlist(temp.df[,jj])), method="linear",na.rm=FALSE))
    # remove NA values that are not confined between numbers with nearest value
    temp.df[,jj]=as.double(na.locf(na.locf(app.var,na.rm=FALSE),na.rm=FALSE, fromLast=TRUE))
  }
  
  #assign the interpolated values into the results list
  temp.name=as.character.Date(dates[ii])
  data1.split[[temp.name]]=data.frame(temp.df)
  rm(temp.df)
}

#remove any empty elemts of the list
data1.split[which(names(data1.split) %in% c(""))]=NULL

summary(data1)

```


## Data interpolation into 1 m Intervals

Use the function "depthinterp"" to interpolate the data into uniform length for every 1 m. This is essential for later calculating the volume of each depth interval and the mass of each N specie measured in this depth.

```{r interpolation, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}

#define matrix size requiered for interpolating the data
n.col.interp=ncol(data1.split[[1]])
n.row.interp=40

#extract unique date values of measurments to be used for real level correction
dates=as.Date(unique(data1$Date, incomparables = FALSE))

# initialize an empty data frame for interpolation results
data1.split.interp=vector("list", length=0)

# run interpolation function for each element
for (ii in 1:length(data1.split)){
  temp=depthinterp(input=na.locf(data1.split[[ii]]))
  temp.name=as.character.Date(dates[ii])
  data1.split.interp[[temp.name]]=data.frame(temp)
}
#remove any empty elemts of the list
data1.split.interp[which(names(data1.split.interp) %in% c(""))]=NULL
```

### Demonstrate interpolations of randomly selected data

```{r interpolation plots, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}
# choose a random number to demonstrate the interpolation
rand.num=sample(1:length(data1.split),1)
rand.date=as.character.Date(dates[rand.num])

plot(data1.split[[rand.date]]$Oxygen,data1.split[[rand.date]]$Depth,  
     main=paste("Oxygen interpolation", rand.date),ylab="Depth [m]", xlab="Oxygen [mol/l]",
     ylim=rev(range(data1.split.interp[[rand.num]]$Depth)),
     xlim=c(0, as.double(max(data1.split[[rand.num]]$Oxygen))))
lines(data1.split.interp[[rand.date]]$Oxygen,data1.split.interp[[rand.date]]$Depth,  col=2)

#place a legend for an oxygen plot
legend(0, 1, c("Measured", "Interpolation"), col = c(1, 2),
       lty = c(NA,1), pch = c(1, NA),
       merge = TRUE, bg = "gray90")

plot(data1.split[[rand.date]]$Nitrate,data1.split[[rand.date]]$Depth,  
     main=paste("N species interpolation", rand.date),ylab="Depth [m]", xlab="Nitrate/NH4 [mol/l]",
     ylim=rev(range(data1.split.interp[[rand.date]]$Depth)), xlim=c(0,1.5*as.double(max(c(range(data1.split[[rand.date]]$Nitrate),range(data1.split[[rand.date]]$NH4))))), col=1)

lines(data1.split.interp[[rand.date]]$Nitrate,data1.split.interp[[rand.date]]$Depth,  col=2)
points(data1.split[[rand.date]]$NH4,data1.split[[rand.date]]$Depth, col=3)
lines(data1.split.interp[[rand.date]]$NH4,data1.split.interp[[rand.date]]$Depth,  col=4)

#place a legend for a geochemical plot
legend(1.05*as.double(max(c(max(data1.split[[rand.num]]$Nitrate),max(data1.split[[rand.num]]$NH4)))), 1, c("Measured NO2", "Interpolation","Measured NH4", "Interpolation"), col = c(1, 2,3,4),
       lty = c(NA,1,NA,1), pch = c(1,NA,1,NA),
       merge = TRUE, bg = "gray90")

```

## Depth-Level Corrections
Calculate the appropriate depth for each measurment using available daily level measurments. Use function "realdepth" and use closest day if no level measurment of that day is available. This correction is essential for later calculation of depth to volume and concentration quantification of N species in the lake.

```{r level-depth correction, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}
# this block arranges the depth of the measurments according to the lake level during that day

# initialize a data frame to accept the corrected data
data1.split.interp.depth=vector("list", length=0)
data1.split.depth=vector("list", length=0)

# run depth correction function for each element
for (ii in 1:length(dates)){
  temp.date=as.character.Date(dates[ii])
  temp=realdepth(depth.date=data3, measurment.date=temp.date, input=data1.split.interp[[temp.date]])
  data1.split.interp.depth[[temp.date]]=data.frame(temp)
  
  temp=realdepth(depth.date=data3, measurment.date=temp.date, input=data1.split[[temp.date]])
  data1.split.depth[[temp.date]]=data.frame(temp)
}
#remove any empty elemets of the list
data1.split.depth[which(names(data1.split.depth) %in% c(""))]=NULL
data1.split.interp.depth[which(names(data1.split.interp.depth) %in% c(""))]=NULL
```

### Demonstrate the results of randomly selected data according to the new depth scale 
```{r level-depth correction plots, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}
# choose a random number to demonstrate the interpolation
rand.num=sample(1:length(data1.split),1)
rand.date=as.character.Date(dates[rand.num])

plot(data1.split.depth[[rand.date]]$Oxygen,data1.split.depth[[rand.date]]$Depth,  
     main=paste("Oxygen interpolation", rand.date),ylab="Depth [m]", xlab="Oxygen [mol/l]",
     ylim=range(data1.split.interp.depth[[rand.num]]$Depth), xlim=c(0, max(as.double(data1.split[[rand.date]]$Oxygen))))
lines(data1.split.interp.depth[[rand.date]]$Oxygen,data1.split.interp.depth[[rand.date]]$Depth,  col=2)

#place a legend for an oxygen plot
legend(0, as.double(max(data1.split.interp.depth[[rand.num]]$Depth)), c("Measured", "Interpolation"), col = c(1, 2),lty = c(NA,1), pch = c(1, NA), merge = TRUE, bg = "gray90")

plot(data1.split.depth[[rand.date]]$Nitrate,data1.split.depth[[rand.date]]$Depth,  
     main=paste("N species interpolation", rand.date),ylab="Depth [m]", xlab="Nitrate/NH4 [mol/l]",
     ylim=range(data1.split.interp.depth[[rand.date]]$Depth), xlim=c(0,1.5*max(c(max(as.double((data1.split[[rand.date]]$Nitrate))),max(as.double(data1.split[[rand.date]]$NH4))))), col=1)
lines(data1.split.interp.depth[[rand.date]]$Nitrate,data1.split.interp.depth[[rand.date]]$Depth,  col=2)
points(data1.split.depth[[rand.date]]$NH4,data1.split.depth[[rand.date]]$Depth, col=3)
lines(data1.split.interp.depth[[rand.date]]$NH4,data1.split.interp.depth[[rand.date]]$Depth,  col=4)

#place a legend for a geochemical plot
legend(1.05*max(c(max(as.double(data1.split[[rand.num]]$Nitrate)),max(as.double(data1.split[[rand.num]]$NH4)))), max(data1.split.interp.depth[[rand.num]]$Depth), c("Measured NO2", "Interpolation","Measured NH4", "Interpolation"), col = c(1, 2,3,4),lty = c(NA,1,NA,1), pch = c(1,NA,1,NA),merge = TRUE, bg = "gray90")

```

## Quantification using lake hypsometric curve

The use of concentrations has limitiations because it may enhance or demise processes that is inhernently "mass-based", and therefore limit our undestanding of the process. In this section the data is quantified, and concentrations are transformed into acmounts via multiplaction of each concentration with its corresponding hypsomentric curve  valueusing function "conctoquant".

```{r quntification, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}
# this block calculates the total amount using the hypsometric curve of the lake

# initialize a data frame to accept the corrected data
data1.split.interp.depth.quant=vector("list", length=0)
data1.split.depth.quant=vector("list", length=0)

# run quantification function for each element
for (ii in 1:length(dates)){
  temp.date=as.character.Date(dates[ii])
  temp=conctoquant(hypso.curve=data2, input=data1.split.interp.depth[[temp.date]])
  data1.split.interp.depth.quant[[temp.date]]=data.frame(temp)
  
  temp=conctoquant(hypso.curve=data2, input=data1.split.depth[[temp.date]])
  data1.split.depth.quant[[temp.date]]=data.frame(temp)
  
}
#remove any empty elemts of the list
data1.split.interp.depth.quant[which(names(data1.split.interp.depth.quant) %in% c(""))]=NULL
data1.split.depth.quant[which(names(data1.split.depth.quant) %in% c(""))]=NULL

km2=xlab=expression(Km^{2})

xlab=paste("Area [",km2,"]", collapse=" ")

plot(data2$SA,data2$WD,main=paste("Lake hypsometric curve"),ylab="Depth [m]", xlab=km2,
     ylim=(range(data2$WD)), xlim=range(data2$SA))

```

### Demonstrate the results according to the new depth scale of randomly selected data
```{r quntification plots, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}
# choose a random number to demonstrate the interpolation
rand.num=sample(1:length(data1.split),1)
rand.date=as.character.Date(dates[rand.num])

plot(data1.split.depth.quant[[rand.date]]$Oxygen,data1.split.depth.quant[[rand.date]]$Depth,  
     main=paste("Oxygen interpolation", rand.date),ylab="Depth [m]", xlab="Oxygen [quant]",
     ylim=range(data1.split.interp.depth[[rand.num]]$Depth), xlim=c(0, max(as.double(data1.split.depth.quant[[rand.date]]$Oxygen))))
lines(data1.split.interp.depth.quant[[rand.date]]$Oxygen,data1.split.interp.depth.quant[[rand.date]]$Depth,  col=2)

#place a legend for an oxygen plot
legend(0, max(as.double(data1.split.interp.depth.quant[[rand.num]]$Depth)), c("Measured", "Interpolation"), col = c(1, 2),lty = c(NA,1), pch = c(1, NA), merge = TRUE, bg = "gray90")

plot(data1.split.depth.quant[[rand.date]]$Nitrate,data1.split.depth.quant[[rand.date]]$Depth,  
     main=paste("N species interpolation", rand.date),ylab="Depth [m]", xlab="Nitrate/NH4 [quant]",
     ylim=range(data1.split.interp.depth[[rand.date]]$Depth), xlim=c(0,1.5*max(c(max(as.double(data1.split.depth.quant[[rand.date]]$Nitrate),max(as.double(data1.split.depth.quant[[rand.date]]$NH4)))))), col=1)

lines(data1.split.interp.depth.quant[[rand.date]]$Nitrate,data1.split.interp.depth.quant[[rand.date]]$Depth,  col=2)

points(data1.split.depth.quant[[rand.date]]$NH4,data1.split.depth.quant[[rand.date]]$Depth, col=3)

lines(data1.split.interp.depth.quant[[rand.date]]$NH4,data1.split.interp.depth.quant[[rand.date]]$Depth,  col=4)

#place a legend for a geochemical plot
legend(1.05*max(c((data1.split.depth.quant[[rand.num]]$Nitrate),max(data1.split.depth.quant[[rand.num]]$NH4))), -215, c("Measured NO2", "Interpolation","Measured NH4", "Interpolation"), col = c(1, 2,3,4),lty = c(NA,1,NA,1), pch = c(1,NA,1,NA),merge = TRUE, bg = "gray90")

```

### save the data into an excel file
```{r quntification plots, echo=FALSE, warning=FALSE, message=FALSE, out.width = '600px', out.height = '300px'}

```
