---
title: "LakeN"
author: "Yoav Ben Dor"
date: "6.12.2016"
output: html_document
---

```{r setup, include=FALSE , root.dir="F:/MyDocuments/GIT/LakeN"}
knitr::opts_chunk$set(echo = FALSE , warning=FALSE)


setwd("F:/MyDocuments/GIT/LakeN/R")
source('depthinterp.R')
source('realdepth.R')
source('conctoquant.R')

setwd("F:/MyDocuments/GIT/LakeN")


```

## Lake N model
# Yoav Ben Dor

This model was written as a project for course 68806 Eco-hydrologic modelling with R
This model illustrates the annual N cycle in a lake in northern Israel, and simulates the transitions from oxydized species into reduced species

# read the data and load relevant libraries
```{r echo=FALSE, warning=FALSE}
# setwd("C:/Users/user/Documents/LakeN")
# load relevant libraries
library(plyr)
library(zoo)
library(dplyr)
library(LakeN)

# read the data for the model
# use the read.table function in the cas of a tab=delimited text file
# use the read.csv function for a , delimited text file
data1=read.table("LakeNData1.txt", header=TRUE)
data2=read.table("LakeNData2.txt", header=TRUE)
data3=read.table("LakeNData3.txt", header=TRUE)
data4=read.table("LakeNData4.txt", header=TRUE)

#split the data set by measurment date
data1.split=split(data1, f=data1$Date)

```

# Interpolate the Data into 1 m Intervals
## Use the function "depthinterp"" to interpolate the data into uniform length for every 1 m
```{r echo=FALSE, warning=FALSE}

#define matrix size requiered for interpolating the data
n.col.interp=7
n.row.interp=40

#extract unique date values of measurments to be used for real level correction
dates=as.Date(unique(data1$Date, incomparables = FALSE))

# initialize an empty data frame for interpolation results
data1.split.interp=vector("list", length(data1.split))

#data1.split.interp=list(matrix(data = NA, ncol=n.col.interp, nrow=n.row.interp))

for (ii in 1:length(data1.split)){
  temp=depthinterp(input=data1.split[[ii]])
  temp.name=as.character.Date(dates[ii])
  data1.split.interp[[temp.name]]=data.frame(temp)
}
#remove any empty elemts of the list
data1.split.interp[which(names(data1.split.interp) %in% c(""))]=NULL
```


## Demonstrate interpolations of randomly selected data
```{r echo=FALSE, warning=FALSE}
# choose a random number to demonstrate the interpolation
rand.num=sample(1:length(data1.split),1)
rand.date=as.character.Date(dates[rand.num])

plot(data1.split[[rand.date]]$Oxygen,data1.split[[rand.date]]$Depth,  
     main=paste("Oxygen interpolation", rand.date),ylab="Depth [m]", xlab="Oxygen [mol/l]",
     ylim=rev(range(data1.split.interp[[rand.num]]$Depth)), xlim=c(0, max(data1.split[[rand.num]]$Oxygen)))
lines(data1.split.interp[[rand.date]]$Oxygen,data1.split.interp[[rand.date]]$Depth,  col=2)

#place a legend for an oxygen plot
legend(0, 1, c("Measured", "Interpolation"), col = c(1, 2),
       lty = c(NA,1), pch = c(1, NA),
       merge = TRUE, bg = "gray90")

plot(data1.split[[rand.date]]$Nitrate,data1.split[[rand.date]]$Depth,  
     main=paste("N species interpolation", rand.date),ylab="Depth [m]", xlab="Nitrate/NH4 [mol/l]",
     ylim=rev(range(data1.split.interp[[rand.date]]$Depth)), xlim=c(0,1.5*max(c(max(data1.split[[rand.date]]$Nitrate),max(data1.split[[rand.date]]$NH4)))), col=1)
lines(data1.split.interp[[rand.date]]$Nitrate,data1.split.interp[[rand.date]]$Depth,  col=2)
points(data1.split[[rand.date]]$NH4,data1.split[[rand.date]]$Depth, col=3)
lines(data1.split.interp[[rand.date]]$NH4,data1.split.interp[[rand.date]]$Depth,  col=4)

#place a legend for a geochemical plot
legend(1.05*max(c(max(data1.split[[rand.num]]$Nitrate),max(data1.split[[rand.num]]$NH4))), 1, c("Measured NO2", "Interpolation","Measured NH4", "Interpolation"), col = c(1, 2,3,4),
       lty = c(NA,1,NA,1), pch = c(1,NA,1,NA),
       merge = TRUE, bg = "gray90")

```

# Depth and Level Corrections
## calculate the appropriate depth for each measurment using available daily level measurments. Use function "realdepth" and use closest day if no level measurment of that day is available.
```{r echo=FALSE, warning=FALSE}
# this block arranges the depth of the measurments according to the lake level during that day

# initialize a data frame to accept the corrected data
data1.split.interp.depth=vector("list", length(data1.split.interp))
data1.split.depth=vector("list", length(data1.split.interp))

for (ii in 1:length(dates)){
  temp.date=as.character.Date(dates[ii])
  temp=realdepth(depth.date=data3, measurment.date=temp.date, input=data1.split.interp[[temp.date]])
  data1.split.interp.depth[[temp.date]]=data.frame(temp)
  
  temp=realdepth(depth.date=data3, measurment.date=temp.date, input=data1.split[[temp.date]])
  data1.split.depth[[temp.date]]=data.frame(temp)
}
#remove any empty elemts of the list
data1.split.depth[which(names(data1.split.depth) %in% c(""))]=NULL
data1.split.interp.depth[which(names(data1.split.interp.depth) %in% c(""))]=NULL
```

# demonstrate the results according to the new depth scale of a random data
```{r echo=FALSE, warning=FALSE}
# choose a random number to demonstrate the interpolation
rand.num=sample(1:length(data1.split),1)
rand.date=as.character.Date(dates[rand.num])

plot(data1.split.depth[[rand.date]]$Oxygen,data1.split.depth[[rand.date]]$Depth,  
     main=paste("Oxygen interpolation", rand.date),ylab="Depth [m]", xlab="Oxygen [mol/l]",
     ylim=range(data1.split.interp.depth[[rand.num]]$Depth), xlim=c(0, max(data1.split[[rand.date]]$Oxygen)))
lines(data1.split.interp.depth[[rand.date]]$Oxygen,data1.split.interp.depth[[rand.date]]$Depth,  col=2)

#place a legend for an oxygen plot
legend(0, max(data1.split.interp.depth[[rand.num]]$Depth), c("Measured", "Interpolation"), col = c(1, 2),lty = c(NA,1), pch = c(1, NA), merge = TRUE, bg = "gray90")

plot(data1.split.depth[[rand.date]]$Nitrate,data1.split.depth[[rand.date]]$Depth,  
     main=paste("N species interpolation", rand.date),ylab="Depth [m]", xlab="Nitrate/NH4 [mol/l]",
     ylim=range(data1.split.interp.depth[[rand.date]]$Depth), xlim=c(0,1.5*max(c(max(data1.split[[rand.date]]$Nitrate),max(data1.split[[rand.date]]$NH4)))), col=1)
lines(data1.split.interp.depth[[rand.date]]$Nitrate,data1.split.interp.depth[[rand.date]]$Depth,  col=2)
points(data1.split.depth[[rand.date]]$NH4,data1.split.depth[[rand.date]]$Depth, col=3)
lines(data1.split.interp.depth[[rand.date]]$NH4,data1.split.interp.depth[[rand.date]]$Depth,  col=4)

#place a legend for a geochemical plot
legend(1.05*max(c(max(data1.split[[rand.num]]$Nitrate),max(data1.split[[rand.num]]$NH4))), max(data1.split.interp.depth[[rand.num]]$Depth), c("Measured NO2", "Interpolation","Measured NH4", "Interpolation"), col = c(1, 2,3,4),lty = c(NA,1,NA,1), pch = c(1,NA,1,NA),merge = TRUE, bg = "gray90")

```

# Quantify the data using lake hypsometric curve

## Quantify the amount of measured species using the hypsomentric curve of the lake using function "conctoquant". Multiply each depth value with its corresponding hypsometric value
```{r echo=FALSE, warning=FALSE}
# this block calculates the total amount using the hypsometric curve of the lake

# initialize a data frame to accept the corrected data
data1.split.interp.depth.quant=vector("list", length(data1.split.interp.depth))
data1.split.depth.quant=vector("list", length(data1.split.depth))

for (ii in 1:length(dates)){
  temp.date=as.character.Date(dates[ii])
  temp=conctoquant(hypso.curve=data2, input=data1.split.interp.depth[[temp.date]])
  data1.split.interp.depth.quant[[temp.date]]=data.frame(temp)
  
  temp=conctoquant(hypso.curve=data2, input=data1.split.depth[[temp.date]])
  data1.split.depth.quant[[temp.date]]=data.frame(temp)
}
#remove any empty elemts of the list
data1.split.interp.depth.quant[which(names(data1.split.interp.depth.quant) %in% c(""))]=NULL
data1.split.depth.quant[which(names(data1.split.depth.quant) %in% c(""))]=NULL
```

## demonstrate the results according to the new depth scale of a random data
```{r echo=FALSE, warning=FALSE}
# choose a random number to demonstrate the interpolation
rand.num=sample(1:length(data1.split),1)
rand.date=as.character.Date(dates[rand.num])

plot(data1.split.depth.quant[[rand.date]]$Oxygen,data1.split.depth.quant[[rand.date]]$Depth,  
     main=paste("Oxygen interpolation", rand.date),ylab="Depth [m]", xlab="Oxygen [quant]",
     ylim=range(data1.split.interp.depth[[rand.num]]$Depth), xlim=c(0, max(data1.split.depth.quant[[rand.date]]$Oxygen)))
lines(data1.split.interp.depth.quant[[rand.date]]$Oxygen,data1.split.interp.depth.quant[[rand.date]]$Depth,  col=2)

#place a legend for an oxygen plot
legend(0, max(data1.split.interp.depth.quant[[rand.num]]$Depth), c("Measured", "Interpolation"), col = c(1, 2),lty = c(NA,1), pch = c(1, NA), merge = TRUE, bg = "gray90")

plot(data1.split.depth.quant[[rand.date]]$Nitrate,data1.split.depth.quant[[rand.date]]$Depth,  
     main=paste("N species interpolation", rand.date),ylab="Depth [m]", xlab="Nitrate/NH4 [quant]",
     ylim=range(data1.split.interp.depth[[rand.date]]$Depth), xlim=c(0,1.5*max(c(max(data1.split.depth.quant[[rand.date]]$Nitrate),max(data1.split.depth.quant[[rand.date]]$NH4)))), col=1)

lines(data1.split.interp.depth.quant[[rand.date]]$Nitrate,data1.split.interp.depth.quant[[rand.date]]$Depth,  col=2)

points(data1.split.depth.quant[[rand.date]]$NH4,data1.split.depth.quant[[rand.date]]$Depth, col=3)

lines(data1.split.interp.depth.quant[[rand.date]]$NH4,data1.split.interp.depth.quant[[rand.date]]$Depth,  col=4)

#place a legend for a geochemical plot
legend(1.05*max(c(max(data1.split.depth.quant[[rand.num]]$Nitrate),max(data1.split.depth.quant[[rand.num]]$NH4))), -215, c("Measured NO2", "Interpolation","Measured NH4", "Interpolation"), col = c(1, 2,3,4),lty = c(NA,1,NA,1), pch = c(1,NA,1,NA),merge = TRUE, bg = "gray90")

```
